An examination of the high-order dynamic interactions underlying multi-dimensional timeseries data

### 2019-03-26 paper/main.tex
Level 1 inter-subject correlation heatmap calculations, redo plots, and WIP paper.

\subsection*{Evaluation metrics}
We evaluate our approach to extracting dynamic correlations and
higher-order correlations using several metrics detailed next.  First,
we generated synthetic data using known time-varying correlations, and
then we evaluated the fidelity with which Equation~\ref{eqn:timecorr}
could recover those correlations (for synthetic datasets with
different properties, and using different kernels to define the
weights; Fig.~\ref{fig:weights}).  We then turned to a series of
analyses on a (real) neuroimaging dataset where the ground truth
correlations were \textit{not} known.  We evaluated whether the
recovered correlations could be used to accurately label held-out
neuroimaging data with the time at which it was collected, and also
whether the history of correlations (and higher-order correlations)
from time 0 to $t-1$ could be used to predict future activity patterns
at time $t$.  We used these latter evaluations (using timepoint
decoding and predictions of held-out future data) as a proxy for
gauging how much explanatory power the recovered correlations held
with respect to the observed data.

@@

\subsubsection*{Recovery of ground truth parameters from synthetic
  data}
Apply timecorr with a given kernel, then correlate each recovered
correlation matrix with the ground truth.  Explore how recovery varies
with the kernel, kernel parameters, and specific structure of the data
(e.g. slow changes as in the ramping dataset, versus rapid changes as
in the block dataset).

@@

To assess decoding accuracy, we randomly divided participants for each stimulus into training and testing groups. For the zeroth order, we computed the mean factor activity for each group.  For all subsequent orders up to the tenth order, we computed the mean approximated dynamic ISFC of factor activity for each group.  For each group of participants in turn, we compared these activity patterns (using Pearson correlations) to estimate the story times each pattern corresponded to.

Specifically, we asked, for each timepoint: what are the correlations
between the first group's and second group's activity patterns at each
order.  We then took weighted mixture of the correlation matrices at
each order using mixing parameter, $\phi$, where $0 <\phi< 1$ reflects a
weighted mixture of order based decoding. We optimized  $\phi$, by subdividing the training group. We note that the decoding test we used is a conservative in which we count a timepoint label as incorrect if it is not an exact match.

@@

\subsubsection*{Neuroimaging dataset~\citep{SimoEtal16}}
Figures: decoding by level
For our decoding analysis, we used HTFA-derived node activities
~\cite{MannEtal18} from fMRI data collected as participants listened
to an audio recording of a story (intact condition; 36 participants),
listened to time scrambled recordings of the same story (17
participants in the paragraph-scrambled condition listened to the
paragraphs in a randomized order and 36 in the word-scrambled
condition listened to the words in a randomized order), or lay resting
with their eyes open in the scanner (rest condition; 36
participants). We sought to demonstrate how higher-order correlations
may be used to examine dynamic interactions of brain patterns in
(real) multi-subject fMRI datasets. This story listening dataset was
collected as part of a separate study, where the full imaging
parameters, image preprocessing methods, and experimental details may
be found~\citep{SimoEtal16}. The dataset is available at
\href{url}{http://arks. princeton.edu/ark:/88435/dsp015d86p269k}.

@@

Separately for each experimental condition, we divided participants
into two groups. For the zeroth order, we computed the mean factor
activity for each group.  For all subsequent orders up to the tenth
order, we computed the mean approximated dynamic ISFC of factor
activity for each group (see \textit{Materials and methods}).  For
each order, we correlated the group 1 activity patterns with group 2
activity patterns.  We then subdivided the group 1 to obtain an
optimal weighting parameter for each order’s correlation matrix using
the same cross validation method. We used the optimal weighting
parameters to obtain a weighted-mixture of each order’s correlation
matrix. Using these correlations, we labeled the group 1 timepoints
using the group 2 timepoints with which they were most highly
correlated; we then computed the proportion of correctly labeled group
1 timepoints. (We also performed the symmetric analysis whereby we
labeled the group 2 timepoints using the group 1 timepoints as a
template.) We repeated this procedure 10 times (randomly re-assigning
participants to the two groups each time) to obtain a distribution of
decoding accuracies for each experimental condition. (There were 272
timepoints for paragraph condition, 300 timepoints for intact and word
conditions, and 400 timepoints for rest condition,  so chance
performance on this decoding test is was $\frac{1}{272}$,
$\frac{1}{300}$, and $\frac{1}{400}$ respectively.

>>>>

\subsection*{Evaluation metrics}
We evaluate our approach to extracting dynamic correlations and
higher-order correlations using several metrics detailed next.  First,
we generated synthetic data using known time-varying correlations, and
then we evaluated the fidelity with which Equation~\ref{eqn:timecorr}
could recover those correlations (for synthetic datasets with
different properties, and using different kernels to define the
weights; Fig.~\ref{fig:weights}).  We then turned to a series of
analyses on a (real) neuroimaging dataset where the ground truth
correlations were \textit{not} known.  We evaluated whether the
recovered correlations could be used to accurately label held-out
neuroimaging data with the time at which it was collected.  We used this latter evaluations (using timepoint
decoding) as a proxy for
gauging how much explanatory power the recovered correlations held
with respect to the observed data.

@@

\subsubsection*{Recovery of ground truth parameters from synthetic
  data}
We applied timecorr with a given kernel Fig.~\ref{fig:weights}) to
synthetic data, then correlate each recovered
correlation matrix with the ground truth.  Explore how recovery varies
with the kernel, kernel parameters, and specific structure of the data
(e.g. slow changes as in the ramping dataset, versus rapid changes as
in the block dataset).

@@

To assess decoding accuracy, we randomly divided participants for each
stimulus into training and testing groups. For the zeroth order, we
computed the mean factor activity for each group.  For all subsequent
orders up to the tenth order, we computed the mean approximated
dynamic ISFC of factor activity for each group. To assess how
additional higher-order correlations contribute to decoding accuracy,
for each order we included a weighted-mixutre (described below) of the activity patterns of all
previous orders.  For each group of participants in turn, we compared these activity patterns (using Pearson correlations) to estimate the story times each pattern corresponded to. Specifically, we asked, for each timepoint: what are the correlations
between the first group's and second group's activity patterns at each
order. We note that the decoding test we used is a conservative in which we count a timepoint label as incorrect if it is not an exact match.

For each order we obtained the weighted-mixture of the correlation
matrices for the current order and all previous orders using mixing parameter , $\phi$, where $0 <\phi< 1$ reflects a
weighted mixture of order based decoding Fig.~\ref{fig:decode_level}
Panel C. ). We calculated  $\phi$, by
subdividing the training group and using the quasi-Newton method of
Broyden, Fletcher, Goldfarb, and Shanno (BFGS) for optimization. We
repeated this cross-validation process 100 times.

@@

\subsubsection*{Neuroimaging dataset~\citep{SimoEtal16}}
For our decoding analysis, we used HTFA-derived node activities
~\cite{MannEtal18} from fMRI data collected as participants listened
to an audio recording of a story (intact condition; 36 participants),
listened to time scrambled recordings of the same story (17
participants in the paragraph-scrambled condition listened to the
paragraphs in a randomized order and 36 in the word-scrambled
condition listened to the words in a randomized order), or lay resting
with their eyes open in the scanner (rest condition; 36
participants). We sought to demonstrate how higher-order correlations
may be used to examine dynamic interactions of brain patterns in
(real) multi-subject fMRI datasets. This story listening dataset was
collected as part of a separate study, where the full imaging
parameters, image preprocessing methods, and experimental details may
be found~\citep{SimoEtal16}. The dataset is available at
\href{url}{http://arks. princeton.edu/ark:/88435/dsp015d86p269k}.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level.pdf}
  \caption{\textbf{Decoding by order.
      a. Relative decoding accuracy by order.} Relative to chance,
    decoding accuracy for each condition (intact, paragraph, word,
    and rest) as a function of increasingly more orders.  The red dots indicates
    maximum decoding accuracy for each condition. \textbf{b. Normalized decoding accuracy
      by order.}  We normalized the decoding accuracy by order to better
    visualize the order with the maximum decoding accuracy for each
    condition. \textbf{c. Optimized weights.}
  For the order with maximum decoding accuracy by condition, we show
  barplots of the optimized weights $\phi$  for each contributing
  order.}
  \label{fig:decoding_level}
\end{figure}

@@

Separately for each experimental condition, we divided participants
into two groups. For the zeroth order, we computed the mean factor
activity for each group.  For all subsequent orders up to the tenth
order, we computed the mean approximated dynamic ISFC of factor
activity for each group (see \textit{Materials and methods}), and
combined in a weighted mixutre with all previous orders
(i.e. cross-validation for the second
order contained a weighted-mixture of zeroth, first, and second order \textbf{c. Optimized weights.}.)  For
each order, we correlated the group 1 activity patterns with group 2
activity patterns.  We then subdivided the group 1 to obtain an
optimal weighting parameter for each order’s correlation matrix using
the same cross validation method. We used the optimal weighting
parameters to obtain a weighted-mixture (see \textit{Materials and methods}) of each order’s correlation
matrix. Using these correlations, we labeled the group 1 timepoints
using the group 2 timepoints with which they were most highly
correlated; we then computed the proportion of correctly labeled group
1 timepoints. (We also performed the symmetric analysis whereby we
labeled the group 2 timepoints using the group 1 timepoints as a
template.) We repeated this procedure 100 times (randomly re-assigning
participants to the two groups each time) to obtain a distribution of
decoding accuracies for each experimental condition. (There were 272
timepoints for paragraph condition, 300 timepoints for intact and word
conditions, and 400 timepoints for rest condition,  so chance
performance on this decoding test is was $\frac{1}{272}$,
$\frac{1}{300}$, and $\frac{1}{400}$ respectively.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/corrmats.pdf}
  \caption{\textbf{Inter-subject correlations by order.
      a. Inter-subject correlation heatmaps by order.}  Heatmap
    denoting how correlated each subject is with every other subject
    for each order, and grouped by condition (intact (I), paragraph
    (P), word (W),
    and rest (R)). \textbf{b. Distribution of correlations for each
      condition by order}
   Inter-subject correlations by conditon for each order.}
  \label{fig:corrmats}
\end{figure}
